Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos

### 2014-05-02 paper/main.tex
new changes

There is a long observational tradition studying galactic properties
as a function of their large scale environment
\citep[e.g.][]{Oemler1974,Dressler1980,Pimbblet2002,Gomez2003,Kauffmann2004,Abbas2006,Baldry2006,Park2007,OMill2008,Gonzalez2009,Padilla2010,Wilman2010}. In
these situations the environment definition is usually based on
quantities easily accessible to observations such as local number
density or nearest neighbour measurements \citep{Muldrew2012}. With
the advent of large surveys the visibility of the cosmic web became
clear. As a consequence, in order to capture its filamentarity, the
environment definition started to be more complex, involving shear
properties of the galaxy density field or the reconstructed dark
matter density field
\citep[e.g.]{Lee2005,Sousbie2008,MunozCuartas2011,Tempel2014}.

@@

{\bf Forero-Romero et al. (2014)} & T-Web & $0.5-1 \hMpc$ &
filament &$++$ & $>10^{12}$\hMsun\\
&   & &
filament & $+$ & $<10^{12}$\hMsun\\

&   & &
wall & $++$ & $>10^{12}$\hMsun\\

&   & &
wall & $+$ & $<10^{12}$\hMsun\\\hline

{\bf Forero-Romero et al. (2014)} & Vp-Web & $0.5-1 \hMpc$ &
filament &$--$ & $>10^{12}$\hMsun\\
&   & &
filament & none & $<10^{12}$\hMsun\\
&   & &
wall & $--$ & $>10^{12}$\hMsun\\

&   & &
wall & none & $<10^{12}$\hMsun\\\hline

@@

{\bf Forero-Romero et al. (2014)} & T-Web & $0.5-1$ &
filament & - & $>10^{12}$\hMsun\\

&   & &
filament & none & $<10^{12}$\hMsun\\

&   & &
wall & none & $>10^{12}$\hMsun\\

&   & &
wall & none & $<10^{12}$\hMsun\\\hline

{\bf Forero-Romero et al. (2014)} & Vp-Web & $0.5-1$ &
filament &none & $>10^{12}$\hMsun\\

&   & &
filament & none & $<10^{12}$\hMsun\\

&   & &
wall & + & $>10^{12}$\hMsun\\

&   & &
wall & nonte & $<10^{12}$\hMsun\\\hline

@@

We note that we measure alignments along the eigenvectors of the
cosmic web without defining first whether each point corresponds to a
filament or a wall. Our motivation is two-fold. First, partitioning the
simulation into web elements implies a choice regarding the value
for the parameter $\lambda_{\rm th}$; although this has been done
before we consider that deriving robust results independent of the
free parameter $\lambda_{\rm th}$ can be more illuminating to
understand the connection of halos with the cosmic web.

However, for simplicity and readability we report our results in terms
of alignment with respect to filaments and sheets. Given that the
direction along a filament should be defined by the eigenvector
$\hat{e}_3$ corresponding to the smallest eigenvalue $\lambda_3$, a
strong alignment along that vector will be reported as an alignment
along filaments. Correspondingly, the first eigenvector $\hat{e}_{1}$
defines the direction perpendicular to walls, a strong alignment along
this vector will be reported as an anti-alignment along walls.

@@

The angular momentum in the Vweb (middle row, left column) presents 
a signal of alignment along the second eigenvector $\hat{e}_{2}$;
between $45\%$ to $60\%$ of the halos are aligned along that
direction, while there is a minority of halos aligned with
$\vec{e}_{1}$.  There is a clear change in trends around
$10^{11}\hMsun-10^{12}\hMsun$ depending on the grid resolution;
below that mass range there is no evidence for alignment while at
higher masses all the trends we describe are apparent. This means that most
of the halos tend to lie along walls but do not have a clear alignment
with respect to filaments.

@@

@@

J.E.F-R acknowledges the financial support by Vicerrector\'ia de
Investigaciones at Universidad de los Andes through a FAPA grant.
NP acknowledges support by Fondecyt Regular No. 1110328 and BASAL PFB-06 "Centro de Astronomia y Tecnologias Afines".
NP and SC acknowledge support by the European Commissions Framework Programme 7, through the Marie Curie International
Research Staff Exchange Scheme LACEGAL (PIRSES-GA-2010-269264).
NP thanks the hospitality of the Max Planck
Institute for Astrophysics at Garching during his sabbatical year 2013-2014.
The Geryon cluster at the Centro de Astro-Ingenieria UC was extensively used for the calculations performed in this paper.
The Anillo ACT-86, FONDEQUIP AIC-57, and QUIMAL 130008 provided funding for several improvements to the Geryon cluster.

@@

They used the Tweb method\footnote{In \citep{Hahn2007} the authors do not use the term "Tweb", but their definition of the
tidal field tensor they use is equivalent with the T-Web} applied on three simulations each of $512^3$
particles, with sizes $L_1=45$\hMpc, $L_2=90$\hMpc and
$L_3=180$\hMpc, this corresponds to particle masses of $4.7, 38.0,
300\times 10^7$\hMsun. Halo identification was done with a FOF
algorithm with 0.2 times the interparticle distance. They considered
halos of at least 300  particles.

>>>>

There is a long observational tradition studying galactic properties
as a function of their large scale environment
\citep[e.g.][]{Oemler1974,Dressler1980,Pimbblet2002,Gomez2003,Kauffmann2004,Abbas2006,Baldry2006,Park2007,OMill2008,Gonzalez2009,Padilla2010,Wilman2010,Muldrew2012}. In
these situations the environment definition is usually based on
quantities easily accessible to observations such as local number
density or nearest neighbour measurements. With
the advent of large surveys the visibility of the cosmic web became
clear. As a consequence, in order to capture its filamentarity, the
environment definition started to be more complex, involving shear and gradient 
properties of the galaxy density field or the reconstructed dark
matter density/velocity field
\citep[e.g.]{Lee2005,AragonCalvo2007,Hahn2007,Sousbie2008,Zhang2009,MunozCuartas2011,Vweb,Trowland2013,Tempel2014}.

@@

{\bf Forero-Romero et al. (2014)} & T-Web & $0.5-1 \hMpc$ &
$\hat{e}_3$ (filament) &$++$ & $>10^{12}$\hMsun\\
&   & &
$\hat{e}_3$ (filament) & $+$ & $<10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & $++$ & $>10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & $+$ & $<10^{12}$\hMsun\\\hline

{\bf Forero-Romero et al. (2014)} & Vp-Web & $0.5-1 \hMpc$ &
$\hat{e}_3$ (filament) &$--$ & $>10^{12}$\hMsun\\
&   & &
$\hat{e}_3$ (filament) & none & $<10^{12}$\hMsun\\
&   & &
$\hat{e}_1$ (wall) & $--$ & $>10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & none & $<10^{12}$\hMsun\\\hline

@@

{\bf Forero-Romero et al. (2014)} & T-Web & $0.5-1$ &
$\hat{e}_3$  (filament) & - & $>10^{12}$\hMsun\\

&   & &
$\hat{e}_3$  (filament) & none & $<10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & none & $>10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & none & $<10^{12}$\hMsun\\\hline

{\bf Forero-Romero et al. (2014)} & Vp-Web & $0.5-1$ &
$\hat{e}_3$  (filament) & none & $>10^{12}$\hMsun\\

&   & &
$\hat{e}_3$ (filament) & none & $<10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & + & $>10^{12}$\hMsun\\

&   & &
$\hat{e}_1$ (wall) & none & $<10^{12}$\hMsun\\\hline

@@

We note that we measure alignments along the eigenvectors of the
cosmic web without defining first whether each point corresponds to a
filament or a wall. However, for simplicity and readability we describe
our results in terms of alignment with respect to filaments and
sheets. Given that the direction along a filament should be defined by the
eigenvector $\hat{e}_3$ corresponding to the smallest eigenvalue $\lambda_3$, a
strong alignment along that vector will be reported as an alignment
along filaments. Correspondingly, the first eigenvector $\hat{e}_{1}$
defines the direction perpendicular to walls, a strong alignment along
this vector will be reported as an anti-alignment along walls.

We avoid the classification into filaments and walls for the following
reason. Partitioning the simulation into web elements implies
a choice regarding the value for the parameter $\lambda_{\rm th}$;
although this has been done before, we consider that deriving robust
results independent of the free parameter $\lambda_{\rm th}$ can be
more illuminating to understand the connection of halos with the
cosmic web.

However, a possible disadvantage is that we are mixing the alignment signal
of different environments. For instance, if half of halo population of
fixed mass is aligned along filaments (and the vector $\hat{e}_3$), and the
other half along planes (perpendicular to $\hat{e}_1$ but without a clear signal along
$\hat{e}_3$, $\hat{e}_2$), the total signal of the alignment along
$e_3$ might appear diluted in comparison to a signal taken separately
for filaments and walls.

@@

The angular momentum in the Vweb (middle row, left column) presents 
a signal of alignment along the second eigenvector $\hat{e}_{2}$;
between $45\%$ to $60\%$ of the halos are aligned along that
direction, while there is a minority of halos aligned with
$\vec{e}_{1}$.  There is a clear change in trends around
$10^{11}\hMsun-10^{12}\hMsun$ depending on the grid resolution;
below that mass range there is no evidence for alignment while at
higher masses all the trends we describe are noticeable. This means
that the angular momentum of halos above $10^{12}\hMsun$ tends to lie
along walls, parallel to the vector $\vec{e}_2$; without any clear
trend for alignment/anti-alignment with respec to filaments.

@@

It is possible to consider that numerical effects can also
affect the estimation of the eigenvectors. For instance,
different levels of shot noise in the interpolation scheme to
construct the density and momentum grids could contribute to the
anti-alignment of the two algorithms. However, a detailed study of the
inter-web alignments is beyond the scope of this paper.

@@

J.E.F-R acknowledges the financial support by Vicerrector\'ia de
Investigaciones at Universidad de los Andes through a FAPA grant.
NP acknowledges support by Fondecyt Regular No. 1110328 and BASAL PFB-06 "Centro de Astronomia y Tecnologias Afines".
NP and SC acknowledge support by the European Commissions Framework
Programme 7, through the Marie Curie International Research Staff
Exchange Scheme LACEGAL (PIRSES-GA-2010-269264). NP thanks the
hospitality of the Max Planck Institute for Astrophysics at Garching
during his sabbatical year 2013-2014. The Geryon cluster at the Centro de Astro-Ingenieria UC was extensively used for the calculations performed in this paper.
The Anillo ACT-86, FONDEQUIP AIC-57, and QUIMAL 130008 provided
funding for several i provements to the Geryon cluster.

We thank the referee, Miguel Aragon-Calvo, for a constructive report
that help us to improve the paper.

@@

They used the hessian of the gravitational potential \footnote{In \citep{Hahn2007} the authors use a definition of the tidal field tensor equivalent with the T-Web method} applied on three simulations each of $512^3$
particles, with sizes $L_1=45$\hMpc, $L_2=90$\hMpc and
$L_3=180$\hMpc, this corresponds to particle masses of $4.7, 38.0,
300\times 10^7$\hMsun. Halo identification was done with a FOF
algorithm with 0.2 times the interparticle distance. They considered
halos of at least 300  particles.
